Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets

### 2021-05-21 epsilon-removal.tex
applicaion partielle transducteur

% fork of main.tex with more expressive models
% SWT read both tapes (under 2 heads) at every transition (even epsilon or semi-epsilon)
% SWA read tape (under head) at every transition (even epsilon)
% SW-VPA read top of stack (when not empty) at every transition

\documentclass[runningheads]{llncs}
%\documentclass[a4paper,11pt]{article}
%\setcounter{page}{1}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{hyperref}
%\usepackage[bookmarks,bookmarksnumbered,naturalnames,plainpages=false]{hyperref}
%usepackage{url}

% for footnote ref
\usepackage{refcount}

% array and tabular
\usepackage{array}
\newcolumntype{C}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

% extension of enumerate env. (style for displaying counters)
% \usepackage{enumerate}

%% pictures
% \usepackage{graphicx} 
% \DeclareGraphicsExtensions{.pdf,.png,.jpg}
% \graphicspath{fig/}

%% PGF, Tikz
\usepackage{tikz-cd}
%% \usepackage{pgfplots}
%% \usepgfplotslibrary{dateplot}
%% \usepackage{pgf,pgfarrows,pgfnodes, pgfautomata}
% \usepackage{tikz}
% \usetikzlibrary{cd}
%% \usetikzlibrary{arrows}
%% \usetikzlibrary{calc}
%% \usetikzlibrary{snakes}
%% \usetikzlibrary{backgrounds}
% \usetikzlibrary{trees}
%% \usetikzlibrary{automata}
%% \usetikzlibrary{positioning}
%% \usetikzlibrary{matrix}
%% \usetikzlibrary{patterns}
%% \usetikzlibrary{shapes}

% symbols
\usepackage{amsmath} 
\usepackage{amssymb} 
\usepackage{amsbsy}
\usepackage{bbold}
\usepackage{latexsym}
%\usepackage{amsfonts}
\usepackage{stmaryrd}
%\usepackage{mathabx}
%\usepackage{MnSymbol}
\usepackage{harmony} % simple music fonts
\usepackage{mathtools} % for arrows
%\usepackage{mathptmx}

%% theorem environments
\usepackage{theorem}
% \newtheorem{theorem}{Theorem} %[section]
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \newtheorem{proposition}[theorem]{Proposition}
% \newenvironment{proof}{\vspace{-2ex}{\it Proof. }}{\hspace*{\fill} $\Box$\smallskip }
% \theorembodyfont{\slshape}
% \newtheorem{example}[theorem]{Example}
% \newtheorem{remark}[theorem]{Remark}
\def\endex{\hspace*{\fill} $\Diamond$\smallskip }
{\theorembodyfont{\rmfamily} \theoremstyle{break} \newtheorem{algo}{Algorithm}}

%% algorithms
%\usepackage{algorithm} 
%\usepackage{program}

%% for new macros
\usepackage{xspace}

%% arrows etc
%\input{rewriting}

%% music symbols
% see http://tug.ctan.org/info/latex4musicians/latex4musicians.pdf
\usepackage{musicography}

%% Misc macros

\def\ie{\textit{i.e.}\xspace}
\def\eg{\textit{e.g.}\xspace}
\def\wrt{\textit{wrt}\xspace}
%\def\wlog{\textit{wlog}\xspace}
\def\etc{\textit{etc}\xspace}

\def\<#1>{\langle #1 \rangle}
\newcommand{\pair}[2]{\langle{#1}, {#2}\rangle}
%\newcommand{\A}{\mathcal{A}}
%\newcommand{\B}{\mathcal{B}}
\newcommand{\D}{\mathbb{D}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\W}{\mathbb{W}}
\newcommand{\T}{\mathcal{T}}

\newcommand{\Semiring}{\mathbb{S}}
\newcommand{\zero}{\mathbb{0}}
\newcommand{\one}{\mathbb{1}}
\newcommand{\dom}{\ensuremath{\mathit{dom}}}

\def\SA{\textsf{sA}\xspace}
\def\WA{\textsf{wA}\xspace}
\def\SWT{\textsf{swT}\xspace}
\def\SWA{\textsf{swA}\xspace}
\def\SWTA{\textsf{swTA}\xspace}
\def\SWVPA{\textsf{sw-VPA}\xspace}
\def\weight{\mathsf{weight}}
\def\wei{\mathsf{w}}
\def\mei{\mathsf{m}}
\def\init{\mathsf{in}}
\def\final{\mathsf{out}}
\newcommand{\call}[1]{\ensuremath #1} %{\ensuremath \langle_{#1}}
\newcommand{\return}[1]{\ensuremath #1} %{\ensuremath {}_{#1}{\rangle}} % $\prescript{}{a}{)}$
\def\Omegai{{\Omega_\mathsf{i}}}
\def\Omegac{{\Omega_\mathsf{c}}}
\def\Omegar{{\Omega_\mathsf{r}}}
\def\Sigmai{{\Sigma_\mathsf{i}}}
\def\Sigmac{{\Sigma_\mathsf{c}}}
\def\Sigmar{{\Sigma_\mathsf{r}}}
\def\Deltai{{\Delta_\mathsf{i}}}
\def\Deltac{{\Delta_\mathsf{c}}}
\def\Deltar{{\Delta_\mathsf{r}}}
\def\Phii{{\Phi_\mathsf{i}}}
\def\Phic{{\Phi_\mathsf{c}}}
\def\Phir{{\Phi_\mathsf{r}}}
\def\Phici{{\Phi_\mathsf{ci}}}
\def\Phicc{{\Phi_\mathsf{cc}}}
\def\Phicr{{\Phi_\mathsf{cr}}}
\def\Phicx{{\Phi_\mathsf{cx}}}
\def\weii{{\wei_\mathsf{i}}}
\def\weic{{\wei_\mathsf{c}}}
\def\weir{{\wei_\mathsf{r}}}
\def\weie{{\wei_\mathsf{e}}}
\newcommand{\config}[2]{\ensuremath \genfrac{[}{]}{0pt}{}{#1}{#2}} 
\newcommand{\opluseq}{\ensuremath\mathrel{\oplus}=}
\newcommand{\ioi}[1]{\mathsf{ioi}({#1})}
\newcommand{\rank}{\mathsf{rk}}
\newcommand{\lin}{\mathsf{lin}}

%\sloppy

% Parsing over infinite alphabet as optimal alignment computation 
% as edit-distance between string and language
% 
%\title{Symbolic Weighted Parsing and Automated Music Transcription}
%\title{Symbolic Weighted Language Models and Automated Music Transcription}
\title{Symbolic Weighted Language Models and\\ Parsing over Infinite Alphabets}
%\title{Weighted Visibly Pushdown Automata and Automated Music Transcription}
\author{Florent Jacquemard}
\institute{INRIA \& CNAM, Paris, France\\
\email{florent.jacquemard@inria.fr}}

%\titlerunning{WVPA \& AMT}
%\titlerunning{Symbolic Weighted Parsing} % and Automated Music Transcription
%\authorrunning{Florent Jacquemard}

\date{\today}

\begin{document}
\thispagestyle{empty}
\maketitle

\begin{abstract}
\input{abstract}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% intro
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction} \label{sec:intro}
\input{introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SWT & SWA
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{SW Automata and Transducers}
\label{section:transducer}\label{sec:transducer}
\label{section:SWA}\label{sec:SWA}
\label{section:SWT}\label{sec:SWT}

We follow the approach of~\cite{Mohri03EDWA} for the computation of distances,
between words and languages, using weighted transducers, 
and extend it to infinite alphabets.
% with models of symbolic weighted automata and transducers. 
%
The models introduced in this section generalize 
weighted automata and transducers~\cite{Droste09handbook} 
%over finite alphabets, see  e.g.~\cite{Mohri03EDWA}, 
by labelling each transition with a weight function (instead of a simple weight value), 
that takes the input and output symbols as parameters. 
These functions are similar to the guards of symbolic automata~\cite{dAntoniVeanes17CAV,dAntoni21CACM},
but they can return values in an generic semiring, 
whereas the latter guards are restricted to the Boolean semiring.

\subsection{Semirings} 
\label{section:semiring}\label{sec:semiring}
\input{semiring}

\subsection{Label Theory}
\label{section:symbols}
\input{labels}

\subsection{Definitions} \label{sec:SWTdef}\label{sec:SWAdef}
%(SWT)

Let $\Semiring$ be a commutative and complete semiring, 
$\Sigma$ and $\Delta$ be alphabets called respectively \emph{input} and \emph{output}, %{alphabets}, 
and $\bar\Phi = \< \Phi_\Sigma, \Phi_\Delta, \Phi_{\Sigma, \Delta}>$ be an $\Semiring$-label theory.

\begin{definition}
\label{def:transducer} \label{def:SWT}
A \emph{symbolic-weighted transducer} (\SWT)
over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$
%the input and output alphabets~$\Sigma$ and $\Delta$ with label theory $\bar\Phi$, and the semiring $\Semiring$ 
is a tuple
$T = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Phi_{\Sigma, \Delta}$, 
respectively $\mathsf{out} : Q \to \Phi_{\Sigma, \Delta}$,
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a 4-uplet of transition functions 
$\wei_{00}$, $\wei_{10}$, $\wei_{01}$, and $\wei_{11}$
%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
from $Q \times Q$ into $\Phi_{\Sigma, \Delta}$.
\end{definition}

%A pair of states $\< q, q'>$ such that $\wei_{00}(q, q') \neq \zero$
%is called an \emph{$\epsilon$-transition}.
%
For convenience, when $\wei_{ij}(q, q') = \eta \in \Phi_{\Sigma, \Delta}$, 
for $q, q' \in Q$ and $i, j \in \{ 0, 1\}$,
respectively $\mathsf{in}(q) = \eta$, $\mathsf{out}(q') = \eta$,
we shall sometimes write, overloading the function names:
$\wei_{ij}(q, a, b, q') = \eta(a, b)$ for $a \in \Sigma$,  $b \in \Delta$,
respectively $\mathsf{in}(q, a, b) = \eta(a, b)$,
$\mathsf{out}(q', a, b) = \eta(a, b)$.

\noindent 
The \SWT $T$ computes on pairs of words $\< s, t> \in \Sigma^+ \times \Delta^+$,
$s$ and $t$, being respectively called input and output word.
%
% \noindent
More precisely, the symbolic-weighted transducer~$T$ defines a mapping 
from the pairs of strings of $\Sigma^+ \times \Delta^+$ 
into~$\Semiring$,
based on the following intermediate function $\weight_T$
defined recursively for every $q, q' \in Q$, 
%$a \in \Sigma$, $b \in \Delta$ 
%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
$a \in \Sigma$, $u \in \Sigma^*$, $b\in \Delta$, $v\in \Delta^*$, 
\[
\begin{array}{lccl}
%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
\weight_T(q, au, bv, q') & = & 
   \displaystyle\bigoplus_{r \in Q} &
   \wei_{00}(q, a, b, r) \otimes \weight_T(r, au, bv, q')\\
  & \oplus & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
    \wei_{10}(q, a, b, r) \otimes \weight_T(r, u, bv, q')\\
  & \oplus & \displaystyle\bigoplus_{r \in Q, v \neq \epsilon} &
    \wei_{01}(q, a, b, r) \otimes \weight_T(r, au, v, q')\\
  & \oplus & \displaystyle\bigoplus_{r \in Q, u,v \neq \epsilon} &
    \wei_{11}(q, a, b, r) \otimes \weight_T(r, u, v, q')\\
  & \oplus & \displaystyle\bigoplus_{u,v = \epsilon} &
    \wei_{11}(q, a, b, q')
\end{array}
\]
%
We recall that, by convention (Section~\ref{sec:semiring}), 
an empty sum with $\bigoplus$ is equal to~$\zero$. 
Since $\zero$ is absorbing for~$\otimes$ in~$\Semiring$,
one term $\wei_{ij}(q, a, b, q'')$ equal to $\zero$ in the above expression 
will be ignored in the sum, meaning that there is no possible transition
from state $q$ into state $q'$ while reading $a$ and $b$.
%(including the case $a = \epsilon$ and $b = \epsilon$).
%These functions $\phi$ act as guards for the transducer's transitions, 
%preventing a transition when they return the absorbing $\zero$ of $\Semiring$.
This is analogous to the case of a transition's guard not satisfied by $\<a, b>$ for 
symbolic transducers.
% 
The above sum may be infinite because of the transitions $\wei_{00}$. % in the first and second lines.
In this case, the infinite sum operators applies since $\Semiring$ is assumed complete.
%As we shall see, under conditions, the $\epsilon$-transitions can be removed and hence the sum made finite.
%** SUPPR. $\epsilon$-transition (with $\wei_{00}$) first. **
%
%*** rem. there is no case for $a, \epsilon$ or $\epsilon, b$.
%   It can be simulated with $\wei_{10}$ and $\wei_{01}$ on the last letter. ***

%** The application of $\oplus$ can be seen as a ND choice between transitions **
% choose to advance head in input, in output or both (or none)
Intuitively, using a transition $\wei_{ij}(q, a, b, q')$ means for $T$:
when reading respectively $a$ and $b$ at the current positions in the input and output words, 
increment the current position in the input word if if $i = 1$ (otherwise, do not change it),
increment the current position in the output word if if $j = 1$, 
and change state from $q$ to $q'$.
%
In contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
the input and output symbols at current positions are read by every transitions, 
even transitions that do not increment the head's position.
%This is in contrast with the models of weighted transducers over finite alphabets~\cite{Mohri03ijfcs},
%which ignore the symbol at current position in this case.
This is an important feature in the case of an infinite alphabet, 
since symbols cannot be stored in the finite memory of the transducer in this case.
%whereas considering the current symbol may be useful to compute a transition weight.
%(even when it does not change the head's position, like with $\epsilon$-transitions).
%
%The cases $\weight_T(q, au, \epsilon, q')$ and $\weight_T(q, \epsilon, bv, q')$
%are missing in the definition of $\weight_T$.
%It means that $T$ must avoid configurations where it reached the end of 
%the output word and not of the input one, or vice-versa.
%This can be done by using $\wei_{10}$ and $\wei_{01}$
%before reaching the end of word, and using a special state for this purpose.

The expression of $\weight_T$ can be seen as a stateful definition of 
an edit-distance between a word $s \in \Sigma^*$ and a word $t \in \Delta^*$,
see also~\cite{Mohri03ijfcs}.
Intuitively, 
$\wei_{10}(q, a, b, r)$ is the cost of 
the deletion of the symbol~$a \in \Sigma$ in~$s$, 
$\wei{01}(q, a, b, r)$ is the cost 
of the insertion of~$b \in \Delta$ in $t$, 
and $\wei_{11}(q, a, b, r)$ is the cost 
of the substitution of  $a \in \Sigma$ by~$b \in \Delta$.
%
The cost of a sequence of such operations transforming $s$ into $t$, 
is the product, with $\otimes$, of the individual costs of the operations involved;
And the distance between $s$ and $t$ is the sum, with $\oplus$,
of all such product of costs.

\medskip\noindent
Let $\< s, t> \in \Sigma^+ \times \Delta^+$, with $s = s_0\ldots s_m$, $t = t_0\ldots t_n$). 
The weight associated by $T$ to $\< s, t>$
is defined as follows:
\begin{equation}
T(s, t)  = 
\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_0, t_0) 
\mathop{\otimes} \weight_T(q, s, t, q') \mathop{\otimes} \mathsf{out}(q', s_m, t_n)
\label{eq:SWT-value}
\end{equation}

\begin{example}
We can define with a \SWT the computation of a similarity measure 
between timed sequences
similar to dynamic time warping (DTW).

Let $\Semiring$  be the tropical (\emph{min-plus}) semiring of Figure~\ref{fig:semirings} and 
let $\Sigma = \Delta = \mathbb{R}_+$ be sets of timestamps.
We consider a \SWT with one state $q$ and transitions
$\wei_{11}(q, d, d', q) = 
 \wei_{01}(q, d, d', q) = 
 \wei_{10}(q, d, d', q) = |d' - d|$,
and $\wei_{00}(q, d, d', q) = \zero$, for all $d, d' \in \mathbb{R}_+$.
% needs reading input/output symbols even by epsilon-transitions.
The recursive definition of $\weight_T$ correspond to the dynamic programming equations of DTW
for the computation of an optimal match between words, 
the matching cost for two symbols being the 
the time distance between them.
\endex
\end{example}

\begin{example}
% simple pointwise distance between two sequences of timestamped events **
Let us consider the tropical (\emph{min-plus}) semiring~$\Semiring$ 
of Figure~\ref{fig:semirings} and 
let $\Sigma = \mathbb{R}_+$ be an input alphabet of event dates
and $\Delta = \{ \mathsf{e}, \mathsf{-} \} \times \mathbb{R}_+$ 
be an output alphabet of symbols with timestamps. 
A symbol $\< \mathsf{e}, d > \in \Delta$ represents an event starting at date $d$, 
and $\< \mathsf{-}, d >$ is a continuation of the previous event.
%More precisely, we let 
%$\Sigma = \{ a, - \} \times \R_+$ and 
This example of $\Delta$ is motivated by the case of music notation, 
where several notated events (notes) can be tied together, 
with a \emph{tie} or a \emph{dot}
(like in $\musQuarter\!\!\mathrel{\raisebox{-1.5mm}{$\smile$}}\!\!\musEighth$ 
 or equivalently~\musQuarterDotted),
meaning that they will be played as a unique sounding event.
%The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

We consider a \SWT with two states $q_0$ and $q_1$ whose purpose 
is to compare a recorded performance $s \in \Sigma^*$
with notated music sheet $t \in \Delta^*$.
One timestamp $d_i \in \Sigma$ may corresponds 
to one notated event $\<\mathsf{e}, d'_i> \in \Sigma$, in which case 
the weight value computed by the \SWT is the time distance between both
(see transitions $\wei_{11}$ below).
%
Moreover, if $\<\mathsf{e}, d'_i>$ is followed by continuations 
$\<\mathsf{-}, d'_{i+1}>$..., they are just skip with no cost (transitions $\wei_{01}$ or weight $\one$).
%transitions
\[
\begin{array}{rclcrcl}
\wei_{11}(q_0, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d| & \quad &
\wei_{11}(q_1, d, \< \mathsf{e}, d'>, q_0) & = & |d' - d|\\
\wei_{01}(q_0, d, \< \mathsf{-}, d'>, q_0) & = & \one & &
\wei_{01}(q_1, d, \< \mathsf{-}, d'>, q_0) & = & \one\\
\wei_{10}(q_0, d, b, q_1) & = & \alpha & & \multicolumn{3}{l}{\mathrm{for~all~} b \in \Delta}
\end{array}
\]
%
Moreover, it may happen that the performers plays an extra note accidentally, but only once in a row. 
This is modelled by the transition $\wei_{10}$ with an arbitrary weight value $\alpha \in \Semiring$, 
switching from state $q_0$ (normal) to $q_1$ (error).
The transitions in the second column below switch back to the normal state $q_0$.
% the metric computed by the \SWT is the smallest sum of point wise distances 
% between dates of input and output events.
At last, we let $q_0$ be the only initial and final state, with
$\init(q_0) = \one$, $\init(q_1) = \zero$, 
$\final(q_0) = \one$, $\final(q_1) = \zero$, 
\endex
\end{example}

\noindent
The \emph{Symbolic Weighted Automata} %$A = \< Q, \init, \weight, \final >$
%over $\Sigma$,  $\Semiring$ and $\bar\Phi$
are defined similarly as the transducers of Definition~\ref{def:SWT}, 
by simply omitting the output symbols.
%
In this case, the label theory $\bar\Phi$ can be reduced to a singleton $\< \Phi_\Sigma>$.
%over $\Sigma$ is reduced to
%a set $\Phi_\Sigma$ closed under~$\oplus$ and~$\otimes$.
%
\begin{definition} \label{def:SWA}
A \emph{symbolic-weighted automaton} (\SWA)
over $\Sigma$, $\Semiring$ and $\bar\Phi$
%the input alphabet~$\Sigma$ and the commutative semiring $\Semiring$ 
is a tuple
$A = \< Q, \init, \bar{\wei}, \final >$,
where $Q$ is a finite set of states, 
$\mathsf{in} : Q \to \Phi_\Sigma$, 
respectively $\mathsf{out} : Q \to \Phi_\Sigma,$
are functions defining the weight for entering, 
respectively leaving, a state, 
and $\bar{\wei}$ is a pair of transition functions 
$\wei_{0}$, $\wei_{1}$
%$\wei_{i, j}: Q \times Q \to \Phi_{\Sigma, \Delta}$, with $i, j \in \{ 0, 1\}$.
from $Q \times Q$ into $\Phi_{\Sigma}$.
\end{definition}
A pair of states $\< q, q'>$ such that $\wei_{0}(q, q') \neq \zero$
is called an \emph{$\epsilon$-transition}.

\noindent
As above in the case of \SWT, 
the transition functions $\wei_0$ and $\wei_1$
can be synthesized into a single function
$\wei: Q \times \Sigma \times Q \to \Semiring$, 
and the computation of $A$ on words $s \in \Sigma^+$
is defined with an intermediate function $\weight_A$, 
%$a \in \Sigma$, $b \in \Delta$ 
%for every strings of $s \in \Sigma^*$, $t \in \Delta^*$:
defined as follows for $q, q' \in Q$, $a \in \Sigma$, $u \in \Sigma^*$,
\[
\begin{array}{lccl}
%\weight_T(q, \epsilon, \epsilon, q)  & = & \multicolumn{2}{l}{\one}\\ % or \final(q)
%\weight_T(q, \epsilon, \epsilon, q') & = & \multicolumn{2}{l}{\zero \mathrm{~if~} q \neq q'}\\
\weight_A(q, au, q') & = & 
   \displaystyle\bigoplus_{r \in Q} &
   \wei_{0}(q, a, r) \otimes \weight_A(r, au, q')\\
  & \oplus & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
    \wei_{1}(q, a, r) \otimes \weight_A(r, u, q')\\
  & \oplus & \displaystyle\bigoplus_{u = \epsilon} & 
    \wei_{1}(q, a, q')
\end{array}
\]
%
\noindent
and the weight value associated by $A$ to 
$s = s_0\ldots s_m \in \Sigma^+$ is: % defined as follows:
\begin{equation}
A(s)  = 
\displaystyle\bigoplus_{q, q' \in Q} \mathsf{in}(q, s_0) 
\mathop{\otimes} \weight_A(q, s, q') \mathop{\otimes} \mathsf{out}(q', s_m)
\label{eq:SWA-value}
\end{equation}

%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

%The \emph{summary} of a $\SWT$, resp. a $\SWA$, is ***

\subsection{Properties}
The two following properties will be useful to our approach on 
symbolic weighted parsing in Section~\ref{sec:parsing}.

\begin{proposition} \label{prop:epsilon}
Given a \SWT $T$ over $\Sigma$, $\Delta$, $\Semiring$ and $\bar\Phi$,
and $s \in \Sigma^*$, 
there exists an effectively constructible \SWA 
$A_{T, s}$ over $\Delta$, $\Semiring$ and $\bar\Phi$,
such that for all $t \in \Delta^*$, $A_{T, s}(t) = T(s, t)$.
\end{proposition}
%
\begin{proof}
Let $T = \< Q, \init, \bar{\wei}, \final >$,
let 
$\wei: Q \times \Sigma \times \Delta \times Q \to \Semiring$
be the synthesized form of 
$\bar{\wei} = \< \wei_{00}, \wei_{10}, \wei_{01}, \wei_{11}>$,
and let $s = s_1 \ldots s_n$ with $s_1, \ldots, s_n \in \Sigma$.
%$n = |s|$ be the length of $s$.
%
The state set of $A_{T, s}$ will be $Q' = [1..n] \times Q$.
Its state entering weight function is defined by 
$\init'(1, q) = \init(q)$ for all $q \in Q$, 
and $\init'(i, q) = \zero$ for all $1 < i \leq n$, 
and its state leaving weight function by 
$\final'(n, q) = \final(q)$ or all $q \in Q$, 
and $\final'(i, q) = \zero$ for all $1 \leq i <  n$.

\noindent
We define transition functions $\wei'_0$ and $\wei'_1$ for $A_{T, s}$, 
together with their synthesized form $\wei'$ as follows: 
for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclcl}
\wei'_0\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{00}(q, q')\\ 
%
\wei'_0\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
\mathrm{if} & i < n, \mathrm{where~} \phi = \wei_{10}(q, q'),\\
%
\wei'_0\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
\mathrm{if} & j \neq i, i+1,\\ 
%
\wei'_1\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_{01}(q, q')\\
%
\wei'_1\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
\mathrm{if} & i < n, \mathrm{where~} \eta = \wei_{11}(q, q'),\\
%
\wei'_1\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
\mathrm{if} & j \neq i, i+1.
\end{array}      
\]
We use the assumption that the underlying label theory of $T$ is 
is closed under partial applications (Section~\ref{section:symbols})
in the above construction of $\wei'_1$.

\noindent
The $\SWA$ wanted is $A_{T, s} = \< Q', \init', {\bar\wei}', \final' >$.
\qed
\end{proof}
%
\noindent
The construction time and size for $A_{T, s}$ are $O(\| T \| . | s |)$,
where the size $\| T \|$ of $T$ is its number of states $|Q|$.

% gen. to transducers
\medskip
The following proposition shows that the transitions of type $\wei_{00}$
(analogous of $\epsilon$-transitions of automata over finite alphabets)
are not necessary to the expressiveness of $\SWA$, 
and can safely nullified, 
with an appropriate update of the other transitions.
%
\begin{proposition}
When $\Semiring$ is commutative, bounded and complete, 
given a \SWA $A$ 
over $\Sigma$, $\Semiring$ and $\bar\Phi$
there exists an effectively constructible \SWA $A'$ 
over $\Sigma$, $\Semiring$ and $\bar\Phi$,
without $\epsilon$-transitions,
such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
\end{proposition}
%
\begin{proof} 
Let $A = \< Q, \init, \<\wei_0, \wei_1>, \final >$.
%
We associate to every sequence in $Q^+$ called \emph{$\epsilon$-path}, 
and every $a \in \Sigma$, a weight value 
defined recursively as follows (for $q_1, q_2 \in Q$ and $\sigma \in Q^*$):
\[
\begin{array}{rcl}
\weight_\epsilon(q_1, a) & = & \one\\
\weight_\epsilon(q_1 q_2 \sigma, a) & = & \wei_0(q_1, a, q_2) \otimes \weight_\epsilon(q_2 \sigma, a)\\
\end{array}
\]
%
We shall construct a \SWA $A' = \< Q, \init, \<\wei'_0, \wei'_1>, \final' >$,
where, for all $q, q' \in Q$, and $a \in \Sigma$, 
and 
\[
\begin{array}{rcl}
\wei'_0(q, a, q') & = & \zero\\
\wei'_1(q, a, q') & = & 
\displaystyle\bigoplus_{r \in Q} 
\displaystyle\bigoplus_{\sigma \in Q^*} 
\weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q')\\
\final'(q, a) & = & 
\final(a, q) \oplus
\displaystyle\bigoplus_{r \in Q} 
\displaystyle\bigoplus_{\sigma \in Q^*} 
\weight_\epsilon(q\sigma r, a) \otimes \final(r, a)\\
\end{array}
\]
Hence $\wei'_0(q, q')$ is the constant function $\zero$ of $\Phi_\Sigma$.
Following Equation~(\ref{eq:SWA-value}),
\[
\begin{array}{lccl}
\weight_{A'}(q, au, q') & = & 
    \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
    \wei'_{1}(q, a, r) \otimes \weight_{A'}(r, u, q')
    \oplus \displaystyle\bigoplus_{u = \epsilon} \wei'_{1}(q, a, q') \\
 & = & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} & 
  \displaystyle\bigoplus_{\sigma \in Q^*} 
  \weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q') \otimes \weight_{A'}(r, u, q')\\
 & \oplus & \displaystyle\bigoplus_{r \in Q, u = \epsilon} &
   \displaystyle\bigoplus_{\sigma \in Q^*} \weight_\epsilon(q\sigma r, a) \otimes \wei_{1}(r, a, q')  
\end{array}
\]
Moreover, we can show that
\[
\begin{array}{lccl}
\weight_{A}(q, au, q') & = & 
    \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
    \wei'_{1}(q, a, r) \otimes \weight_{A'}(r, u, q')\\
 & = & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} & 
  \displaystyle\bigoplus_{\sigma \in Q^*} 
  \weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q')
\end{array}
\]

>>>>

% gen. to transducers
\medskip
The following proposition shows that the transitions of type $\wei_{00}$
(analogous of $\epsilon$-transitions of automata over finite alphabets)
are not necessary to the expressiveness of $\SWA$, 
and can safely nullified, 
with an appropriate update of the other transitions.
%
\begin{proposition}
When $\Semiring$ is commutative, bounded and complete, 
given a \SWA $A$ 
over $\Sigma$, $\Semiring$ and $\bar\Phi$
there exists an effectively constructible \SWA $A'$ 
over $\Sigma$, $\Semiring$ and $\bar\Phi$,
without $\epsilon$-transitions,
such that for all $s \in \Sigma^*$, $A'(s) = A(s)$.
\end{proposition}
%
\begin{proof} 
Let $A = \< Q, \init, \<\wei_0, \wei_1>, \final >$.
%
We associate to every sequence in $Q^+$ called \emph{$\epsilon$-path}, 
and every $a \in \Sigma$, a weight value 
defined recursively as follows (for $q_1, q_2 \in Q$ and $\sigma \in Q^*$):
\[
\begin{array}{rcl}
\weight_\epsilon(q_1, a) & = & \one\\
\weight_\epsilon(q_1 q_2 \sigma, a) & = & \wei_0(q_1, a, q_2) \otimes \weight_\epsilon(q_2 \sigma, a)\\
\end{array}
\]
%
We shall construct a \SWA $A' = \< Q, \init, \<\wei'_0, \wei'_1>, \final' >$,
where, for all $q, q' \in Q$, and $a \in \Sigma$, 
and 
\[
\begin{array}{rcl}
\wei'_0(q, a, q') & = & \zero\\
\wei'_1(q, a, q') & = & 
\displaystyle\bigoplus_{r \in Q} 
\displaystyle\bigoplus_{\sigma \in Q^*} 
\weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q')\\
\final'(q, a) & = & 
\final(a, q) \oplus
\displaystyle\bigoplus_{r \in Q} 
\displaystyle\bigoplus_{\sigma \in Q^*} 
\weight_\epsilon(q\sigma r, a) \otimes \final(r, a)\\
\end{array}
\]
Hence $\wei'_0(q, q')$ is the constant function $\zero$ of $\Phi_\Sigma$.
Following Equation~(\ref{eq:SWA-value}),
\[
\begin{array}{lccl}
\weight_{A'}(q, au, q') & = & 
    \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
    \wei'_{1}(q, a, r) \otimes \weight_{A'}(r, u, q')
    \oplus \displaystyle\bigoplus_{u = \epsilon} \wei'_{1}(q, a, q') \\
 & = & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} & 
  \displaystyle\bigoplus_{\sigma \in Q^*} 
  \weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q') \otimes \weight_{A'}(r, u, q')\\
 & \oplus & \displaystyle\bigoplus_{r \in Q, u = \epsilon} &
   \displaystyle\bigoplus_{\sigma \in Q^*} \weight_\epsilon(q\sigma r, a) \otimes \wei_{1}(r, a, q')  
\end{array}
\]
Moreover, %we can show that
\[
\begin{array}{lccl}
\weight_{A}(q, au, q') & = & 
    \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} &
    \wei'_{1}(q, a, r) \otimes \weight_{A'}(r, u, q')\\
 & = & \displaystyle\bigoplus_{r \in Q, u \neq \epsilon} & 
  \displaystyle\bigoplus_{\sigma \in Q^*} 
  \weight_\epsilon(q\sigma r, a) \otimes  \wei_1(r, a, q')
\end{array}
\]
